---
title: "Lesson 4 practices"
author: "S.Omar"
date: "8/28/2020"
output: html_document
---

 In this problem set, you'll continue to explore the diamonds data set.
Your first task is to create a scatterplot of price vs x. using the ggplot syntax.
This assignment is not graded and will be marked as correct when you submit.

ENTER YOUR CODE BELOW THIS LINE
===========================================
```{r}

library(ggplot2)
data("diamonds")
df=diamonds
names(df)
```



```{r}
ggplot(aes(price, x), data=df)+geom_point()
```
### correlation between price vs x:
```{r}
with(df, cor.test(price,x))
```

### correlation between price vs y

```{r}
with(df, cor.test(price,y))
```
### correlation between price vs y

```{r}
with(df, cor.test(price,z))
```



### Create a simple scatter plot of price vs depth.

 ENTER YOUR CODE BELOW THIS LINE
==================================================


```{r}
ggplot(aes(y=price,x= depth), data=df)+geom_point(alpha=0.1/10, position='jitter')
```


###or 
```{r}
ggplot(aes(y = price, x = depth), data = diamonds)+
  geom_jitter(alpha = 0.05)
```

###Adjustment - price VS depth
Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units. See the instructor notesfor two hints.


```{r}
range(df$price)
```


```{r}
ggplot(aes(y= price, x = depth), data = df)+
  geom_jitter(alpha = 1/100)+
  scale_alpha_continuous(breaks=seq(0,18823,2))
```
based on the scatterplot; most diamonds are between the prince range of:(326,6000) and depth range of (59,64)

```{r}
summary(df$price)
```
```{r}
summary(df$depth)
```
###correlaton price vs depth

```{r}
with(df, cor.test(price, depth))
```
NOte: they are rarely corelated.
Q: What is the correlation of depth vs price?

A: -0.01

Q: Based upon the correlation coefficient, would you use depth to predict the price of a diamond?

A: No.

Why?

Because there is not a strong correlation between those variables and what is there is negative.

### Create a scatterplot of price vs carat
and omit the top 1% of price and carat values.

quantile(diamonds$price, probs = c(0.1, 0.99))
     10%      99% 
  646.00 17378.22 

quantile(diamonds$carat, probs = c(0.1, 0.99))
 10%  99% 
0.31 2.18 


ENTER YOUR CODE BELOW THIS LINE
 ========================================

```{r}
quantile(df$carat, probs=c(0.99))
```
```{r}
quantile(df$price, probs=c(0.99))
```

normal plot:
```{r}
p1= ggplot(aes(carat,price), data=df)+geom_point()+
  ggtitle('Normal plot Proce Vs carat')
  

p2= ggplot(aes(carat, price), data=df, 
           xlim=c(0, qunatile(df$carat, 0.99)),
           ylim=c(0,qunatile(df$price,0.99)))+
  geom_point(alpha = 1/100)+
  ggtitle('Price Vs Carat 99% Precentile ')


library(gridExtra)
grid.arrange(p1,p2,ncol=2)  
  
#conventional way using Quantile  
#p2= ggplot(aes(x = carat, y = price), 
#      data = subset(df, 
#                   price < 17378.22 &
#                    carat < 2.18))+
# geom_point(alpha = 1/100)
  
  
```

## what is the maximum size for carat?

```{r}
max(df$carat)
```
## what is maximum size for Price?
```{r}
max(df$price)
```
### Question Template

Q: Does the x-axis for carat extend past 2? It should!

A: Yes it does.


## Create a scatterplot of price vs. volume (x * y * z). This is a very rough approximation for a diamond's volume.
 Create a new variable for volume in the diamonds data frame. This will be useful in a later exercise.Don't make any adjustments to the plot just yet.

 ENTER YOUR CODE BELOW THIS LINE
 =================================================================


```{r}
df$volume=df$x * df$y * df$z
```


```{r}
ggplot(aes(x=volume, y=price), data=df)+geom_jitter(alpha=1/10)
```
## Findings - price vs. volume

What were your observations from the price vs. volume scatterplot?

Response:The distribution was a very dense, tightly bound plot dominated by a couple of outliers.

** Grader Notes: **

Did you notice some outliers? Some volumes are 0!

There’s an expensive diamond with a volume near 4000*, and a cheaper diamond with a volume near 900.

You can find out how many diamonds have 0 volume by using count(diamonds$volume == 0), the count() function comes with the plyr package.

Note: If you ran the count function from plyr, you need to run this command in R to unload the plyr package. detach(“package:plyr”, unload=TRUE)

The plyr package will conflict with the dplyr package in later exercises.

Depending on your investigation, it may or may not be important for you to understand how outliers, like these, came to be in your data.

```{r}

#usig table
print(table(df$volume==0))
```





## what is corr of price and volume.(0<volume<=800)?

```{r}
volume_cor=subset(df, volume<=800 & volume>0)
with(volume_cor, cor(volume, price))
                  
```

lets see the difference:
```{r}
with(df, cor(volume, price))
```



### Subset the data: 
subset the data In order to exclude diamonds with a volume  greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot. (See the Instructor Notes or look up the documentation of geom_smooth() for more details about smoothers.)

Do you think this would be a useful model to estimate the price of diamonds? Why or why not?


 ENTER YOUR CODE BELOW THIS LINE
 ========================================


```{r}
ggplot(aes(y=volume, x= price), data=volume_cor)+geom_point(alpha=1/30)+
  geom_smooth(method='lm')
```
we can see the correlation between two variables and we can see that it will be useful in indicating price prediction. learn more about [smooth modeling methods](https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/)




## Experimenting smoothing model

```{r}
p1 <- ggplot(data = volume_cor,aes(x = volume, y = price)) +geom_point()+

# Default smoother
p2 <- p1 + geom_smooth()+ggtitle("default")

# looking at a linear fit,
p3 <- p1 + stat_smooth(method = "lm", formula = y ~ x, size = 1) + coord_cartesian(ylim = c(0,20000))+
  ggtitle("one degree")

# Looking at polynimoal functions of order 2
p4 <- p1 + stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1) + coord_cartesian(ylim = c(0,20000))+ggtitle("two degree")

# Looking at polynimoal functions of order 3
p5 <- p1 + stat_smooth(method = "lm", formula = y ~ poly(x, 3), size = 1) + coord_cartesian(ylim = c(0,20000))+ggtitle("three degree")


grid.arrange(p2,p3,p4,p5,ncol =2)
```
## Question

Do you think this would be a useful model to estimate the price of diamonds? Why or why not?

Does the linear model seem to be a good fit to the data? Share your thoughts.

Response:

This would be a useful model because the line appears to be pretty good fit to the data.


# Use the function dplyr package to create a new data frame containing info on diamonds by clarity.

Name the data frame diamondsByClarity. The data frame should contain the following variables in this order.
    (1) mean_price
    (2) median_price
    (3) min_price
    (4) max_price
    (5) n
where n is the number of diamonds in each level of clarity.

DO NOT ALTER THE NEXT THREE LINES OF CODE.
 ======================================================
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
data(diamonds)

ENTER YOUR CODE BELOW THIS LINE
 ======================================================

```{r}

library('dplyr')

daimonds_by_clarity= df%>% 
  group_by(clarity)%>% 
  summarise(mean_price=mean(price),
            median_price=median(price),
            min_price=min(price),
            max_price=max(price),
            n=n())%>%
  arrange(clarity)

head(daimonds_by_clarity)
  
  

```

### or in concentional way
```{r}
diamond_Clarity <- group_by(diamonds, clarity)
daimonds_by_clarity <- summarise(diamond_Clarity,
                               mean_price = mean(price),
                               median_price = median(price),
                               min_price = min(price),
                               max_price = max(price),
                               n = n())


head(daimonds_by_clarity)

```


## by clarity and color. 
You can run the code in R to verify what data is in the variables diamonds_mp_by_clarity and diamonds_mp_by_color. Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.


# DO NOT DELETE THE LINES OF CODE BELOW
 ===================================================================
data(diamonds)
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

ENTER YOUR CODE BELOW THIS LINE
===================================================================




```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
diamonds_mp_by_clarity

```



```{r}
diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
diamonds_mp_by_color

```
```{r}
diamonds_by_cut= group_by(df, cut)%>%summarise(mean_price= mean(price))
diamonds_by_cut
```




### Plot
```{r}

plot1 <- ggplot(aes(x = clarity, y = mean_price), 
                data = diamonds_mp_by_clarity)+
  geom_bar(stat = "identity")


plot2 <- ggplot(aes(x = color, y = mean_price),
                data = diamonds_mp_by_color)+
  geom_bar(stat = "identity")

grid.arrange(plot1, plot2, ncol=1)
```



### Add fill command
```{r}

# Mean Price across Clarity and Color

p1  <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price, fill= clarity)) +
  geom_bar(stat = "identity")


p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price, fill= color)) +
  geom_bar(stat = "identity")


grid.arrange(p1,p2, ncol =2)
```
```{r}

p3  <- ggplot(diamonds_by_cut, aes(x = cut, y = mean_price, fill= cut)) +
  geom_bar(stat = "identity")


grid.arrange(p1,p2, p3, ncol = 3)
```

Note: on average calarity does not affect the price of diamonds that much, whic on the other hand level of color can aboslutely affet the price of diamonds.

## Questions

What do you notice in each of the bar charts for mean price by clarity and mean price by color?

Response:

SI2 has the best mean price as in the Clarity group as does J in the color group. Both plots seem to be slightly skewed.
Graders Coments:

We think something odd is going here. These trends seem to go against our intuition.

Mean price tends to decrease as clarity improves. The same can be said for color.

We encourage you to look into the mean price across cut.

UPDATE

The Cut bar chart seems to show some oddities as well …. nothing is correlating with common sense intuition.












